Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-19 latex/article.tex
octree ok

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Curvature Tensor Estimation}
\label{sec:curv-tens-estim}

>>>>

\section{Curvature Tensor Estimation}
\label{sec:curv-tens-estim}

\section{Isosurface Extraction on GPU}
\label{sec:isos-extr-gpu}

In this section, we detail the adaptive isosurface extraction
algorithm. The proposed approach is based on an octree representation
of the input object on which an efficient adaptive triangulation
of the isosurface is constructed on GPU. Such approach is motivated by
the fact that the hierarchical representation of the object allows us
to handle large datasets and to locally adapt the level of details
with respect to the geometry or camera position. We first present the
octree representation and then the isosurface extraction from
hierarchical octree cells.

\subsection{Linear Octree Representation}

\begin{figure}[!htbp]
  \begin{center}
    \subfigure{}{\includegraphics[width=\textwidth]{figs/morton}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/partitioning}}
    %%\subfigure{}{\includegraphics[height=0.17\textheight]{figs/quadtree_activefront}}
  \end{center}
\caption{Morton codes associated to cells of a linear quatree. Each
   morton code of a child cell is obtained by adding a suffix to its
  parent code \emph{(left)}. The adaptive representation consists of
  quadtree cells whose depth is view point dependent
  \emph{(middle)}. Finally, adaptive Marching-Cubes is used to
  generate the triangulation \emph{(right)}.}
\label{fig_quadtree_partitionning}
\end{figure}

Representing a hierachical structure on GPU is usually challenging
since the GPU memory is limited and non consistent \todo{fix coherent/aligned/coalesced?}
memory access (\emph{e.g.} tree traversal using pointer dereferencing)
may lead to huge payload from cache misses.  Efficient spatial tree
encoding can be achieved using pointerless strictures. We use here
linear quadtrees or octrees to represent classical quadtree/octreee
trees, see for instance Gargantini
\cite{gargantini1982effective}. This structure consists in indexing
each cell by a \emph{Morton code}: the code of children cells are
defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). For a given cell, its
code encodes its position with respect to its parent cell and its
complete path to the tree root. Hence, whatever their depth, all cell
codes are simply stored in a linear vector on memory. Furthermore,
cell operations such as subdivision, merging, neighbors access can be
efficiently implemented using bitwise operations on the Morton code.
In the following, we use the GPU implementation proposed by Dupuy
\textit{et al.}  \cite{dupuy2014quadtrees}. In such approach, Morton
codes and bitwise operations are optimized to match with GPU hardware
requirements.

\subsection{Data parallel and adaptive mesh generation}

Using this spatial datastructure, triangulated mesh can be constructed
from a Marching-Cubes approach \cite{lorensen1987marching} (MC for short): the
triangulation is generated from local triangle patches computed on
local cell configurations. Since triangle patches are locally
computed, such approach is fully parallel and easy to implement on GPU
hardwares. However, since adjacent cells may not have the same depth
on the octree, original Lorensen and Cline's rules need to be
updated (see Figure~\ref{fig_quadtree_partitionning}-\emph{right}). Many authors have addressed this problem both for primal and
dual meshes
\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.

In the following, we use the extension of MC configurations to handle
adaptive structures proposed by Lengyel \emph{et al.}
\cite{lengyel2010voxel}. First this approach constrains the octree
structure in order to make sure that the depth difference between any
two adjacent cells is at most one. Then, Lengyel \emph{et al.} propose
specific MC configurations for such transition cells,  leading to a
crack-free triangulation.

Similarly to original MC algorithm, this approach is perfectly adapted
to GPU hardwares: Given a set of cells (a vector of morton codes),
each triangle patch can be constructed in parallel (done by 
in the \emph{geometry shader} in the GPU graphic pipeline).

\subsection{LoD Criteria and Temporal Updates}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{figs/pipeline}
  \caption{ This figure summarizes our GPU pipeline. Data buffers are
    represented in green and computations in red. Each computation
    retrieves data from a buffer and fills a new one. }
  \label{fig_pipeline} 
\end{figure}

\begin{figure}
\centering
\includegraphics[height=0.2\textheight]{subdivision}
\caption{The octree at frame $t$ is stored as a list of Morton codes in a buffer. 
To update the tree, we fill a new buffer representing its state at frame $t+1$. 
On the GPU, each code of the first buffer is processed in parallel. 
If the cell is to be subdivided, it writes the codes of its children in the new buffer. 
If it merges, it writes either nothing, or the code of its parent (if it is the first child).
If it is kept, it just re-writes its own code. }
\label{subdivision}
\end{figure}

\section{Interactive Computation on GPU}
\label{sec:inter-visu-gpu}
